---
title: "01_SPINS_data_cleaning"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r results='hide', message=F, warning=F}
library(dplyr) 

# read in final SPINS data
spins_df <- read.csv("/projects/loliver/SPINS_PLS_Conn/data/raw/SPINSDataStorage_DATA_2020-10-16_1224.csv",stringsAsFactors = F)

```

```{r}
# demographics data cleaning
spins_df$site <- spins_df$redcap_data_access_group
spins_df$diagnostic_group <- ifelse(spins_df$redcap_event_name == 'control_arm_1', 'control', 'case')
spins_df$demo_sex <- ifelse(spins_df$demo_sex_birth == 1, 'female', 'male')

# no one endorsed demo_current_sex_different (though there are some NAs)
spins_df[spins_df$demo_current_sex_different==1 & !is.na(spins_df$demo_current_sex_different),]

# create scanner variable
spins_df$scanner <- gsub("_....$", "", gsub("SPN01_", "", spins_df$record_id))
```

```{r}
# change unknown/missing values to NAs
spins_df$demo_ethnicity[spins_df$demo_ethnicity == 3] <- NA # ethnicity
spins_df$demo_marital_status[spins_df$demo_marital_status == 6|spins_df$demo_marital_status == 9] <- NA # marital status
spins_df$demo_eng_prim_lang[spins_df$demo_eng_prim_lang == 9] <- NA # eng primary lang
spins_df$demo_siblings[spins_df$demo_siblings == 9] <- NA # siblings
spins_df$demo_offspring[spins_df$demo_offspring == 9] <- NA # offspring
spins_df$demo_spec_ed_self[spins_df$demo_spec_ed_self == 9] <- NA # special education
spins_df$demo_highest_grade_self[spins_df$demo_highest_grade_self == 101] <- NA # education
spins_df$demo_highest_grade_mom[spins_df$demo_highest_grade_mom == 101] <- NA # education mom
spins_df$demo_highest_grade_dad[spins_df$demo_highest_grade_dad == 101] <- NA # education dad

```

```{r}
# write out csv with termination explanations
term_info <- spins_df[!is.na(spins_df$term_early_withdraw),c("record_id","diagnostic_group","term_early_withdraw","term_consent_withdraw","term_no_longer_elg","term_change_health","term_deceased_date","term_psych_withdraw","term_other_specify")]

# write.csv(term_info, '/projects/loliver/SPINS_PLS_Conn/data/processed/spins_termination_info_11-20-2020.csv', row.names = FALSE)

```

```{r}
# social and non-social cognition data cleaning

# collapse across sites for WTAR and soc cog variables of interest
spins_df[spins_df$scanner=='MRC'|spins_df$scanner=='MRP',c('wtar_raw_score','wtar_std_score','scog_rmet_total','scog_rad_total',
     'scog_iri_factor_perspective','scog_iri_factor_fantasy','scog_iri_factor_empathic_concern','scog_iri_factor_personal_distress',
     'scog_er40_crt_columnqcrt_value','scog_er40_cr_columnpcr_value','scog_er40ang_columnv','scog_er40fear_columnw',
     'scog_er40hap_columnx','scog_er40noe_columny','scog_er40sad_columnz','scog_tasit_p1_total','scog_tasit_p1_total_positive',
     'scog_tasit_p1_total_negative','scog_tasit_p2_total','scog_tasit_p2_sin','scog_tasit_p2_sscar','scog_tasit_p2_psar',
     'scog_tasit_p2_do','scog_tasit_p2_say','scog_tasit_p2_think','scog_tasit_p2_feel','scog_tasit_p3_total',
     'scog_tasit_p3_lie','scog_tasit_p3_sar','scog_tasit_p3_do','scog_tasit_p3_say','scog_tasit_p3_think','scog_tasit_p3_feel')]  <- 
     spins_df[spins_df$scanner=='MRC'|spins_df$scanner=='MRP',c('wtarsum_total','wtarsum_std','scog2_rmet_total','scog2_rad_total',
     'scog2_iri_factor_perspective','scog2_iri_factor_fantasy','scog2_iri_factor_empathic_concern','scog2_iri_factor_personal_distress',
     'scog2_er40_crt_columnqcrt_value','scog_er402_cr_columnpcr_value','scog2_er40ang_columnv','scog2_er40fear_columnw',
     'scog2_er40hap_columnx','scog2_er40noe_columny','scog2_er40sad_columnz','scog2_tasit_p1_total','scog2_tasit_p1_total_positive',
     'scog2_tasit_p1_total_negative','scog2_tasit_p2_total','scog2_tasit_p2_sin','scog2_tasit_p2_sscar','scog2_tasit_p2_psar',
     'scog2_tasit_p2_do','scog2_tasit_p2_say','scog2_tasit_p2_think','scog2_tasit_p2_feel','scog2_tasit_p3_total',
     'scog2_tasit_p3_lie','scog2_tasit_p3_sar','scog2_tasit_p3_do','scog2_tasit_p3_say','scog2_tasit_p3_think','scog2_tasit_p3_feel')]

# calculate IRI total scores
spins_df$scog_iri_total <- rowSums(spins_df[,c('scog_iri_factor_perspective', 'scog_iri_factor_fantasy', 'scog_iri_factor_empathic_concern', 'scog_iri_factor_personal_distress')], na.rm=TRUE)

# change IRI total 0s to NAs (checked no actual 0s)
spins_df[spins_df$scog_iri_total==0,"scog_iri_total"] <- NA

# reverse score ER-40 RT correct
spins_df$scog_er40_crt <- spins_df$scog_er40_crt_columnqcrt_value*-1 

```

```{r}
# rename variables
spins_df <- spins_df %>% dplyr::rename(scog_iri_perspective = scog_iri_factor_perspective,
  scog_iri_fantasy = scog_iri_factor_fantasy,
  scog_iri_empathic_concern = scog_iri_factor_empathic_concern,
  scog_iri_personal_distress = scog_iri_factor_personal_distress,
  scog_er40_total = scog_er40_cr_columnpcr_value,
  scog_er40_anger = scog_er40ang_columnv,
  scog_er40_fear = scog_er40fear_columnw,
  scog_er40_hap = scog_er40hap_columnx,
  scog_er40_neutral = scog_er40noe_columny,
  scog_er40_sad = scog_er40sad_columnz,
  scog_tasit1_total = scog_tasit_p1_total,
  scog_tasit1_pos = scog_tasit_p1_total_positive,
  scog_tasit1_neg = scog_tasit_p1_total_negative,
  scog_tasit2_total = scog_tasit_p2_total,
  scog_tasit2_sinc = scog_tasit_p2_sin,
  scog_tasit2_simpsar = scog_tasit_p2_sscar,
  scog_tasit2_parsar = scog_tasit_p2_psar,
  scog_tasit2_do = scog_tasit_p2_do,
  scog_tasit2_say = scog_tasit_p2_say,
  scog_tasit2_think = scog_tasit_p2_think,
  scog_tasit2_feel = scog_tasit_p2_feel,
  scog_tasit3_total = scog_tasit_p3_total,
  scog_tasit3_lie = scog_tasit_p3_lie,
  scog_tasit3_sar = scog_tasit_p3_sar,
  scog_tasit3_do = scog_tasit_p3_do,
  scog_tasit3_say = scog_tasit_p3_say,
  scog_tasit3_think = scog_tasit_p3_think,
  scog_tasit3_feel = scog_tasit_p3_feel)

```

```{r results='hide'}
# Clinical and functional data cleaning

# recode SSD disorder type
as.character(spins_df$scid5_dis1_cat_scz_psyc)

spins_df$scid5_SSD_disorder_type <- recode(spins_df$scid5_dis1_cat_scz_psyc, 
  '1' = "Schizophrenia",
  '2' = "Schizophreniform Disorder",
  '3' = "Schizoaffective Disorder",
  '4' = "Delusional Disorder",
  '5' = "Brief Psychotic Disorder",
  '6' = "Psychotic Disorder due to GMC",
  '7' = "Substance-induced Psychotic Disorder",
  '8' = "Psychotic Disorder NOS")
spins_df[spins_df$diagnostic_group=="control","scid5_SSD_disorder_type"] <- NA

# collapse across columns for those with and without regular employment for BSFS
spins_df$bsfs_total <- rowSums(spins_df[,c('bsfs_sec_grandtotal1', 'bsfs_sec_grandtotal2')], na.rm=TRUE)
spins_df[spins_df$bsfs_total == 0,'bsfs_total'] <- NA  # safe as there are no zeros in the original columns

# SANS 2-factor solution as recommended by Dr. Buchanan (email March 6, 2017)
# ref: Blanchard, J. J., & Cohen, A. S. (2005). The structure of negative symptoms within schizophrenia: implications for assessment. Schizophrenia bulletin, 32(2), 238-245.
# ref: Strauss, G. P., Horan, W. P., Kirkpatrick, B., Fischer, B. A., Keller, W. R., Miski, P., Buchanan, R.W., Green, M.F., Carpenter, W. T. (2013). Deconstructing negative symptoms of schizophrenia. Journal of psychiatric research,47(6), 783-790.
# Strauss suggests using the average, not the sum, when calculating the factor scores (excluding globals due to redundancy) because the 2 factors have different numbers of items, which artificially creates a significant within subject effect when you use the sum.

# calculate new factor scores
spins_df$sans_dim_exp <- rowSums(spins_df[,c('sans1_unchanging_expression','sans2_decreased_spont_mvt','sans3_expressive_gestures',
  'sans4_poor_eye','sans5_non_responsivity','sans6_lack_inflection','sans8_poverty_speech','sans10_blocking','sans11_latency')], 
  na.rm=TRUE) * NA^(rowSums(!is.na(spins_df[,c('sans1_unchanging_expression','sans2_decreased_spont_mvt','sans3_expressive_gestures',
  'sans4_poor_eye','sans5_non_responsivity','sans6_lack_inflection','sans8_poverty_speech','sans10_blocking','sans11_latency')]))==0)

# only using sans15_current_role_out as all outpatients (vs sans15b_current_role_in)
spins_df$sans_dim_mot <- rowSums(spins_df[,c('sans13_grooming','sans15a_current_role_out','sans14_role_fn','sans16_anergia',
  'sans19_anhedonia','sans20_sexual_interest','sans21_intimacy','sans18_asociality')], na.rm=TRUE) *  
  NA^(rowSums(!is.na(spins_df[,c('sans13_grooming','sans15a_current_role_out','sans14_role_fn','sans16_anergia',
  'sans19_anhedonia','sans20_sexual_interest','sans21_intimacy','sans18_asociality')]))==0)

# turn factor scores into averages
spins_df$sans_dim_exp_avg <- spins_df$sans_dim_exp/9 
spins_df$sans_dim_mot_avg <- spins_df$sans_dim_mot/8
spins_df[spins_df$diagnostic_group=="control","sans_dim_exp_avg"] <- NA
spins_df[spins_df$diagnostic_group=="control","sans_dim_mot_avg"] <- NA

```

```{r}
# make new df with variables of interest

spins_df <- spins_df[,c("record_id","site","scanner","diagnostic_group","scid5_SSD_disorder_type","demo_sex","demo_age_study_entry",
            "hand_laterality_quotient","demo_marital_status","demo_eng_prim_lang","demo_spec_ed_self","demo_highest_grade_self",
            "term_early_withdraw","wtar_raw_score","wtar_std_score","scog_rmet_total","scog_er40_total","scog_er40_crt",
            "scog_er40_anger","scog_er40_fear","scog_er40_hap","scog_er40_neutral","scog_er40_sad",
            "scog_tasit1_total","scog_tasit1_pos","scog_tasit1_neg","scog_tasit2_total","scog_tasit2_sinc","scog_tasit2_simpsar","scog_tasit2_parsar",
            "scog_tasit2_do","scog_tasit2_say","scog_tasit2_think","scog_tasit2_feel","scog_tasit3_total","scog_tasit3_lie","scog_tasit3_sar",
            "scog_tasit3_do","scog_tasit3_say","scog_tasit3_think","scog_tasit3_feel","scog_rad_total",
            "scog_iri_total","scog_iri_perspective","scog_iri_fantasy","scog_iri_empathic_concern","scog_iri_personal_distress",
            "np_composite_tscore","np_domain_tscore_process_speed","np_domain_tscore_att_vigilance","np_domain_tscore_work_mem",
            "np_domain_tscore_verbal_learning","np_domain_tscore_visual_learning","np_domain_tscore_reasoning_ps",
            "np_domain_tscore_social_cog","np_tscore_tmt","np_tscore_bacsc","np_tscore_hvltr","np_tscore_wms_ss","np_tscore_umd_lns",
            "np_tscore_nab","np_tscore_bvmtr","np_tscore_fluency_animals","np_tscore_cpt","np_tscore_msceit",
            "bsfs_total","bsfs_sec1_total","bsfs_sec2_total","bsfs_sec3_total","bsfs_sec4_total","bsfs_sec5_total",
            "bsfs_sec6_total","bsfs_sec7_y_total_7a","bsfs_sec7_n_total_7b","qls_total","qls20_empathy",
            "qls_factor_interpersonal","qls_factor_instrumental_role","qls_factor_intrapsychic","qls_factor_comm_obj_activities",
            "bprs_factor_total","bprs_factor_neg_symp","bprs_factor_pos_symp","bprs_factor_anxiety_depression","bprs_factor_activation",
            "bprs_factor_hostility","sans_total_sc","sans_dim_mot_avg","sans_dim_exp_avg","sans_sub_affective_flat_blunt",
            "sans_sub_alogia","sans_sub_avolition_apathy","sans_sub_asocial_anhedonia")]

```

```{r}
# exclude participants based on early termination info # N=557

# participants to exclude: 
# term_early_withdraw = 2 (Baseline screening started, no longer eligible) N=40 
# No instances of 4 (Change in physical health), 5 (Hospitalized and unavailable), or 7 (Patient deceased)
ineligible_exc <- term_info[term_info$term_early_withdraw==2,"record_id"]
spins_df <- spins_df[!(spins_df$record_id %in% ineligible_exc),] # N=517


# term_early_withdraw = 1 (Patient withdrew consent) N=26
early_withdraw <- term_info[term_info$term_early_withdraw==1,]
spins_early_withdraw <- spins_df[spins_df$term_early_withdraw==1 & !is.na(spins_df$term_early_withdraw),]

# participants to exclude (N=12): 
# SPN01_CMH_0088 - Participant no longer wished to come in - no data (no visits) - eligibility checklist not completed
# SPN01_CMH_0089 - Too busy with school - no data (no visits) - eligibility checklist not completed
# SPN01_CMH_0102 - dropped out as decided didn't want to do MRI (only did visit 1 - no cog data) - eligibility checklist not completed
# SPN01_CMH_0115 - wary of MRI and substance eligibility unable to be reassessed after positive
# SPN01_CMH_0121 - Participant was not able to abstain for a clean UDS. Decided not to continue in study.
# SPN01_CMH_0127 - found research too taxing (only did visit 1 - no cog data) - dropped out prior to determining eligibility
# SPN01_CMH_0141 - no longer wanted to pursue research - eligibility checklist not completed
# SPN01_CMH_0204 - too busy - eligibility not confirmed (didn't redo UDS)
# SPN01_MRC_0022 - not comfortable disclosing PHI - eligibility checklist not completed
# SPN01_MRP_0111 - changed mind - eligibility checklist not completed
# SPN01_MRP_0126 - no time - screening and eligibility checklist not completed
# SPN01_ZHH_0046 - no longer wanted to participate - not eligible - could not provide informed consent

early_withdraw_exc <- c("SPN01_CMH_0088","SPN01_CMH_0089","SPN01_CMH_0102","SPN01_CMH_0115","SPN01_CMH_0121","SPN01_CMH_0127",
                        "SPN01_CMH_0141","SPN01_CMH_0204","SPN01_MRC_0022","SPN01_MRP_0111","SPN01_MRP_0126","SPN01_ZHH_0046")
spins_df <- spins_df[!(spins_df$record_id %in% early_withdraw_exc),] # N=505

# participants to include:
  # no cog data
# SPN01_CMH_0082 - dropped out as decided didn't want to do MRI (only did visit 1 - no cog or MRI data) - participant never marked eligible re metal implants (x ray was needed), but otherwise met criteria
# SPN01_CMH_0108 - decided didn't want to continue (did visit 1 but no QLS or BSFS - no cog or MRI data)
# SPN01_CMH_0122 - no longer had the time (only did visit 1 - no cog or MRI data)
# SPN01_CMP_0173 - used cannabis over weekend and decided to withdraw after (UDS negative before; only did visit 1 - no cog or MRI data)
# SPN01_CMP_0202 - didn't want to take time off and didn't like MRI (did Visits 1 and 2 - MRI, but no cog data)
# SPN01_MRP_0090 - dropped out as didn't want to get x-ray pre MRI (only did visit 1 - no cog or MRI data)
# SPN01_MRP_0130 - no time (only did visit 1 - no cog or MRI data)
# SPN01_ZHH_0016 - claustrophobic and stressed by study (only did visit 1 - no cog or MRI data)
# SPN01_ZHH_0050 - stressed by study (did visit 1, but no QLS or BSFS - no cog or MRI data)
# SPN01_ZHP_0075 - withdrew pre-MRI (only did visit 1 - no cog or MRI data)
# SPN01_ZHP_0080 - withdrew after MRI, but no MRI data in archive (noted EA and ImObs data not usable; completed visits 1 and 2 - no cog or MRI data)
# SPN01_ZHP_0162 - didn't want to continue (did Visits 1 and 2 - MRI, but no cog data)
  # cog data
# SPN01_CMH_0189 - new job (did visits 1 and 3 - cog data aside from the ER40 (no explanation as to why), but no MRI)
# SPN01_ZHH_0026 - withdrew in scanner (did visits 1 and 3, but no MRI data)


# term_early_withdraw = 8 (Withdrawn by study psychiatrist) N=15
withdrawn <- term_info[term_info$term_early_withdraw==8,]
spins_withdrawn <- spins_df[spins_df$term_early_withdraw==8 & !is.na(spins_df$term_early_withdraw),]

# participants to exclude (N=11 + 1 categorized as could not tolerate MRI, below):
# SPN01_CMH_0009 - high hbA1c values - DC
# SPN01_CMH_0035 - med changes - eligibility checklist not completed
# SPN01_CMH_0037 - clinically unstable - didn't even complete the SCID
# SPN01_CMH_0060 - too large for MRI - eligibility checklist not completed
# SPN01_CMH_0092 - multiple positive UDS - not eligible
# SPN01_CMH_0099 - LOC > 30 mins - eligibility checklist not completed
# SPN01_CMH_0116 - determined mid screening not eligible 
# SPN01_CMH_0132 - high hbA1c values - not eligible
# SPN01_CMH_0163 - positive UDS and unreliable reporting - not eligible
# SPN01_CMH_0210 - determined not to have psychosis by clinical team - diagnosis discrepancy - ineligible 
# SPN01_CMP_0178 - positive UDS (determined late and couldn't be reached to redo) - not eligible

psych_withdraw_exc <- c("SPN01_CMH_0009","SPN01_CMH_0035","SPN01_CMH_0037","SPN01_CMH_0060","SPN01_CMH_0092","SPN01_CMH_0099",
                        "SPN01_CMH_0116","SPN01_CMH_0132","SPN01_CMH_0163","SPN01_CMH_0210","SPN01_CMP_0178")
spins_df <- spins_df[!(spins_df$record_id %in% psych_withdraw_exc),] # N=494

# participants to include:
  # no cog data
# SPN01_CMH_0032 - movement medication side effects (only did visit 1 - no cog or MRI data)
# SPN01_CMH_0081 - withdrew pre-MRI (only did visit 1 - no cog or MRI data)
# SPN01_CMP_0174 - determined to be clinically unstable (only did visit 1 - no cog or MRI data)
# SPN01_CMP_0208 - missed too many MRIs (only did visit 1 - no cog or MRI data)


# explore those who remain from other categories (N=22):
exp_term <- spins_df[spins_df$term_early_withdraw!=2 & spins_df$term_early_withdraw!=1 & spins_df$term_early_withdraw!=8 & !is.na(spins_df$term_early_withdraw),]
exp_term <- exp_term[order(exp_term$term_early_withdraw),] 

# term_early_withdraw = 3 (Unable to contact) N=6; 2 excluded
# participants to exclude:
# SPN01_CMH_0019 - eligibility checklist not completed
# SPN01_MRP_0108 - eligibility checklist not completed

# participants to include:
  # no cog data
# SPN01_CMP_0183 - unable to contact (did Visits 1 and 2 - MRI, but no cog data)
# SPN01_MRC_0059 - unable to contact (only did visit 1 - no cog or MRI data)
# SPN01_ZHP_0103 - unable to contact (did Visits 1 and 2 - MRI, but no cog data)
  # cog data
# SPN01_MRP_0163 - unable to contact (did visits 1 and 3 - cog data, but no MRI data)

# term_early_withdraw = 6 (Could not tolerate MRI) N=8; 1 excluded 
# participants to exclude:
# SPN01_ZHH_0059 - ineligible based on clinician decision (no for able to participant in and provide informed consent) # coding them as excluded due to Withdrawn by Psychiatrist (so as not to confuse with those only excluded from imaging analyses) 

# participants to include: 
  # no cog data
# SPN01_CMH_0010 - withdrew from study after 1st visit - anxious to do MRI (only did visit 1 - no cog or MRI data)
# SPN01_CMH_0045 - could not tolerate MRI (only completed visit 1 - no cog or MRI data)
# SPN01_CMH_0161 - was unable to enter MRI (only completed visit 1 - no cog or MRI data)
# SPN01_MRP_0129 - no notes (only completed visit 1 - no cog or MRI data)
# SPN01_MRP_0131 - MRI coil too small for head (only completed visit 1 - no cog or MRI data)
# SPN01_ZHP_0092 - refused to complete MRI (only completed visit 1 - no cog or MRI data)
  # cog data
# SPN01_CMH_0018 - deemed too large to fit in scanner (did visits 1 and 3 - cog data, but no MRI data)

# term_early_withdraw = 9 (Patient relocated/unavailable for appointments) N=2; 1 excluded
# participants to exclude:
# SPN01_MRP_0141 - unable to obtain UDS (missed appt) - unable to meet eligibility criteria 

# participants to include:
  # no cog data
# SPN01_MRP_0147 - participant moved (did visit 1 but no QLS or BSFS and part of visit 2 (MRI incomplete - RS but no EA, ImObs, or DWI) - no cog data)

# term_early_withdraw = 10 (Other N=6); 2 excluded
# participants to exclude: 
# SPN01_CMH_0064 - participant relocated and required rescreen for positive UDS - not eligible
# SPN01_ZHH_0036 - High HbA1c - not eligible

# participants to include: 
  # no cog data
# SPN01_MRP_0138 - termination checklist says eligible (all answers reflect eligibility) but RA answered no - notes say could not fit in scanner and included in RDOC upload, so I think okay to use existing data (only completed visit 1 - no cog or MRI data) 
# SPN01_MRP_0149 - terminated scan early due to subject not following prompts - moved throughout scan, did not observe/imitate properly, and didn't complete EA (did visits 1 and 2 (imaging probably not usable) - no cog data)
# SPN01_ZHP_0100 - lost to follow up (did visits 1 and 2 - MRI, but no cog data)
# SPN01_ZHP_0135 - lost to follow up (only did visit 1 - no cog or MRI data)

rm_other <- c("SPN01_CMH_0019","SPN01_MRP_0108","SPN01_ZHH_0059","SPN01_MRP_0141","SPN01_CMH_0064","SPN01_ZHH_0036")
spins_df <- spins_df[!(spins_df$record_id %in% rm_other),] # N=488

# CMH_0155 found to have an ASD diagnosis after the end of the study (via SPINS-ASD) - exclude as per Stephanie's recommendation
spins_df <- spins_df[spins_df$record_id!="SPN01_CMH_0155",] # N=487

# one participant completed the study twice, at two different sites - SPN01_ZHP_0130 is the same as SPN01_CMH_0135 - using ZHP_0130 for prisma data (soc cog scores and functioning very similar; cog a bit different, but not all better on re-test)
spins_df <- spins_df[spins_df$record_id!="SPN01_CMH_0135",] # N=486

```

```{r}
# participants missing all cog data (soc and neuro) - recode to NAs (N=30)
no_cog <- c("SPN01_CMH_0082","SPN01_CMH_0108","SPN01_CMH_0122","SPN01_CMP_0173","SPN01_CMP_0202",
            "SPN01_MRP_0090","SPN01_MRP_0130","SPN01_ZHH_0016","SPN01_ZHH_0050","SPN01_ZHP_0075",
            "SPN01_ZHP_0080","SPN01_ZHP_0162","SPN01_CMH_0032","SPN01_CMH_0081","SPN01_CMP_0174",
            "SPN01_CMP_0208","SPN01_CMP_0183","SPN01_MRC_0059","SPN01_ZHP_0103","SPN01_CMH_0010",
            "SPN01_CMH_0045","SPN01_CMH_0161","SPN01_MRP_0129","SPN01_MRP_0131","SPN01_ZHP_0092",
            "SPN01_MRP_0147","SPN01_MRP_0138","SPN01_MRP_0149","SPN01_ZHP_0100","SPN01_ZHP_0135")

spins_df[spins_df$record_id %in% no_cog,16:65] <- NA

# check for other missing cog data
spins_df_cog_check <- spins_df[!(spins_df$record_id %in% no_cog),c(1,16:65,4)]
missing_cog <- spins_df_cog_check[!complete.cases(spins_df_cog_check),c(1:4,10,13:16,21:23,28:29,34:51)] # N=22 (largely missing 1 or 2 measures)

# participants missing all soc cog data - recode to NAs
# SPN01_ZHP_0164 - unable to complete social cog due to cognitive impairment and neg symptoms (but did complete MATRICS); also refused to begin/finish MRI
no_soc_cog <- c("SPN01_ZHP_0164")  
spins_df[spins_df$record_id %in% no_soc_cog,16:47] <- NA

# ZHP_0110 and ZHP_0172 did not complete ER40, RMET, and tasit (only tasit coded incorrectly); ZHP_0172 also did not complete RAD (coded as NA) - change tasit to NAs
spins_df[spins_df$record_id=="SPN01_ZHP_0110" | spins_df$record_id=="SPN01_ZHP_0172",24:41] <- NA

# SPN01_ZHP_0144 refused to complete TASIT after Part 2 - change Part 3 to NAs
# SPN01_ZHP_0140 refused to complete TASIT after Part 3 Q8 - change Part 3 to NAs
# SPN01_ZHP_0094 refused to complete TASIT after Part 3 Q4 - change Part 3 to NAs
# SPN01_CMH_0105 score of 0 for tasit2_simsar total appears to be real
spins_df[spins_df$record_id=="SPN01_ZHP_0144"|spins_df$record_id=="SPN01_ZHP_0140"|spins_df$record_id=="SPN01_ZHP_0094",35:41] <- NA

```

```{r}
# change IDs for participants with repeated scans (using prisma scans in these cases with behavioural from 1st visit)
# SPN01_CMP_0211 has their behavioural data under SPN01_CMH_0211, so change to CMP_0211
spins_df[spins_df$record_id=="SPN01_ZHH_0060","record_id"] <- "SPN01_ZHP_0060"
spins_df[spins_df$record_id=="SPN01_CMH_0180","record_id"] <- "SPN01_CMP_0180"
spins_df[spins_df$record_id=="SPN01_CMH_0182","record_id"] <- "SPN01_CMP_0182"
spins_df[spins_df$record_id=="SPN01_CMH_0191","record_id"] <- "SPN01_CMP_0191"
spins_df[spins_df$record_id=="SPN01_CMH_0196","record_id"] <- "SPN01_CMP_0196"
spins_df[spins_df$record_id=="SPN01_CMH_0198","record_id"] <- "SPN01_CMP_0198"
spins_df[spins_df$record_id=="SPN01_CMH_0207","record_id"] <- "SPN01_CMP_0207"
spins_df[spins_df$record_id=="SPN01_CMH_0211","record_id"] <- "SPN01_CMP_0211" 
spins_df[spins_df$record_id=="SPN01_CMH_0213","record_id"] <- "SPN01_CMP_0213"

# change scanner variable accordingly
spins_df[spins_df$record_id=="SPN01_ZHP_0060","scanner"] <- "ZHP"
CMP <- c("SPN01_CMP_0180","SPN01_CMP_0182","SPN01_CMP_0191","SPN01_CMP_0196","SPN01_CMP_0198",
            "SPN01_CMP_0207","SPN01_CMP_0211","SPN01_CMP_0213")
spins_df[spins_df$record_id %in% CMP,"scanner"] <- "CMP"

# write out inclusion list after early term exclusion
#write.table(spins_df[,"record_id"], '/projects/loliver/SPINS/data/spins_inclusion_09-11-2023.csv', col.names=F, row.names=F)

```

```{r}
library(psych)
library(splitstackshape)

# add mean EA to df
# find EA regressor files
setwd("/scratch/loliver/SPINS_PLS_Conn")
files_ea <- list.files(path= ".", recursive=T, full.names=F, pattern="^.*_EA\\.1D$")

# create list of IDs
ea_ptlist <- paste("SPN01", substring(files_ea,5,7), substring(files_ea,8,11), sep = "_")

# read in ea regressor files
ea <- lapply(files_ea, read.csv, header=F, sep="*")
ea <- lapply(ea, cSplit, splitCols=2:4, sep=":")
names(ea) <- ea_ptlist

# calculate mean EA and add to spins_behav_conn
spins_df$scog_mean_ea <- NA

for (i in names(ea)) {
  spins_df[spins_df$record_id==i,"scog_mean_ea"] <- mean(c(ea[[i]][[1,"V2_1"]],ea[[i]][[1,"V3_1"]],ea[[i]][[1,"V4_1"]],
    ea[[i]][[2,"V2_1"]],ea[[i]][[2,"V3_1"]],ea[[i]][[2,"V4_1"]],ea[[i]][[3,"V2_1"]],ea[[i]][[3,"V3_1"]],ea[[i]][[3,"V4_1"]]))
}

# fisher z transform EA values
spins_df$scog_mean_ea <- fisherz(spins_df$scog_mean_ea)

spins_df <- spins_df[,c(1:41,94,42:93)]

```

```{r}
# organize race and ethnicity data


```

```{r}
# write out a new csv - full sample N=486
#write.csv(spins_df, '/projects/loliver/SPINS_PLS_Conn/data/processed/spins_behav_data_full_09-12-2023.csv', row.names = FALSE)

# write out csv excluding participants with no cog data - N=456
spins_df_cog <- spins_df[!(spins_df$record_id %in% no_cog),]

#write.csv(spins_df_cog, '/projects/loliver/SPINS_PLS_Conn/data/processed/spins_behav_data_cog_09-12-2023.csv', row.names = FALSE)

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
